---
title: "Peak Bloom Prediction by Linear Model"
author: "Hoyong Jung, Taehyoung Kim"
date: "02/13/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Abstract

In this analysis, our group predicts peak bloom date over the next decade for given 4 locations by using multiple linear regression.

## Dependencies

```{r, message=FALSE}
library(tidyverse)
library(rnoaa)
```

## Load the data

The data for each of the four main sites were loaded by tidyverse function called `read_csv` instead of `read.csv`.

```{r, message=FALSE, warning=FALSE}
cherry <- read_csv("data/washingtondc.csv") %>% 
  bind_rows(read_csv("data/liestal.csv")) %>% 
  bind_rows(read_csv("data/vancouver.csv")) %>% 
  bind_rows(read_csv("data/kyoto.csv")) %>%
  mutate(location = factor(location))
```

## Approach

There is a fluctuation over time due to periodic phenomenon like La Niña. (Hence, time series can improve the statistical model.) However, in this analysis, I want to approximate this phenomenon with trigonometric transform based linear regression model.

```{r}
# cherry_fit <- lm(bloom_doy ~ location+lat+long+alt+year, cherry)

cherry_trans <- cherry %>% mutate(
  year_sin = sin(pi / 5.5 * year),
  year_cos = cos(pi / 5.5 * year),
  lat_sin = sin(lat * pi / 180),
  lat_cos = cos(lat * pi / 180),
  long_sin = sin(long * pi / 180),
  long_cos = cos(long * pi / 180),
) %>% 
  select(-bloom_date) %>%
  filter(year > 1990)
  
  

cherry_trans_fit <- lm(bloom_doy ~ ., cherry_trans)

modelbyAIC<-step(cherry_trans_fit,direction="both")
```

## Prediction

```{r}
make_prediction <- function(new_location, new_lat, new_long, new_alt, new_year) {
  new_data <- tibble(year = new_year, year_sin = sin(pi / 5.5 * year), year_cos = cos(pi / 5.5 * year), location = new_location, lat = new_lat, long = new_long, alt = new_alt, 
  lat_sin = sin(lat * pi / 180),
  lat_cos = cos(lat * pi / 180),
  long_sin = sin(long * pi / 180),
  long_cos = cos(long * pi / 180),)
  predict(modelbyAIC,
          newdata=new_data,
          interval = "none",
  )
  
}

predict_tb <- tibble(
  year = c(2023:2032)
)

predict_tb <- predict_tb %>%
  mutate(kyoto = as.integer(make_prediction("kyoto",35.0119831,135.6761135,44, year))) %>%
  mutate(liestal = as.integer(make_prediction("liestal",47.4814,7.730519,350, year))) %>%
  mutate(washingtondc = as.integer(make_prediction("washingtondc",38.8853496,-77.0386278,0, year))) %>%
  mutate(vancouver = as.integer(make_prediction("vancouver",49.2237,-123.1636,24, year)))

predict_tb
```

## Export Data

```{r}
write_csv(predict_tb, "cherry-predictions.csv")
```